---
title: "CelinaScott-Buechler_A1"
author: "Celina Scott-Buechler"
date: "9/23/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

First, I need to install and load necessary packages.

```{r}
#install.packages(c('readr','tidyverse', 'dplyr','zoo', 'plotly', 'sf', 'tigris', 'leaflet', 'censusapi', 'ggplot2', 'ellipsis', 'generics'), dependencies = TRUE)

library('readr')
library('tidyverse')
library('plotly')
library('sf')
library('tigris')
library('leaflet')
library('censusapi')
library('ggplot2')
library('lubridate')
library('zoo')
library('dplyr')
```

Then, I run two loops: 
One that goes through each of the gas use files downloaded from PG&E,
and the other the goes through the electricity use files.

In each case, the loop ones and reads the files, then binds them together 
for a unified data frame. 


```{r, error=TRUE}

years <- 2017:2021
quarters <- 1:4
type <- "Electric"

pge_elec_full <- NULL

for(quarter in quarters) {
  for(year in years){
  
    filename <- 
      paste0(
        "pge/PGE_",
        year,
        "_Q",
        quarter,
        "_",
        type,
        "UsageByZip.csv"
      )
    
    #print(filename)
    
    temp <- read_csv(filename)
    
    pge_elec_full <- rbind(pge_elec_full,temp)
    # Note rbind requires field names to be consistent for every new thing that you add.
    
    saveRDS(pge_elec_full, "pge_elec_full.rds")
}}

pge_elec_full['Total_kBTUs'] <- pge_elec_full$TOTALKWH *3.412

View(pge_elec_full)

```
```{r, error=TRUE}

###Gas

gas <- "Gas"

pge_gas_full <- NULL

for(quarter in quarters) {
  for(year in years){
    
    filename <- 
      paste0(
        "pge/PGE_",
        year,
        "_Q",
        quarter,
        "_",
        gas,
        "UsageByZip.csv"
      )
    
    #print(filename)
    
    temp <- read_csv(filename)
    
    pge_gas_full <- rbind(pge_gas_full,temp)
    # Note rbind requires field names to be consistent for every new thing that you add.
    
    saveRDS(pge_gas_full, "pge_gas_full.rds")
  }}


pge_gas_full['Total_kBTUs'] <- pge_gas_full$TOTALTHM*99976.1/1000

View(pge_gas_full)
```
Then, I filter out only the categories of interest: residential and commercial
consumption of gas and electricity. The month and year columns are combined into
a single date column. Then I can begin graphing. The first graph shows 
residential energy use, broken down by gas and electricity, each month from 2017
to 2021. The next shows the same for commercial energy use.
```{r}

pge_elec_filter <- select(pge_elec_full, c(-TOTALKWH, -AVERAGEKWH)) %>%
  filter(CUSTOMERCLASS %in% c("Elec- Residential","Elec- Commercial"))

#View(pge_elec_filter)


pge_gas_filter <- select(pge_gas_full, c(-TOTALTHM, -AVERAGETHM)) %>%
  filter(CUSTOMERCLASS %in% c("Gas- Residential","Gas- Commercial"))

#View(pge_gas_filter)

pge_both <- rbind(pge_gas_filter, pge_elec_filter)

pge_both$Date <- as.yearmon(paste(pge_both$YEAR, pge_both$MONTH), "%Y %m")
View(pge_both)

pge_both %>% 
  filter(CUSTOMERCLASS %in% c('Gas- Residential','Elec- Residential')) %>%
  ggplot(aes(y=Total_kBTUs, x=Date, fill= CUSTOMERCLASS))+
  geom_bar(stat='identity')
```

```{r pressure, echo=FALSE}
pge_both %>% 
  filter(CUSTOMERCLASS %in% c('Gas- Commercial','Elec- Commercial')) %>%
  ggplot(aes(y=Total_kBTUs, x=Date, fill= CUSTOMERCLASS))+
  geom_bar(stat='identity')

```
To see what effects the pandemic might have had on energy consumption, 
```{r}
pge_both %>% 
  filter(CUSTOMERCLASS %in% c('Gas- Commercial','Elec- Commercial','Gas- Residential','Elec- Residential'), YEAR %in% c(2019,2020,2021)) %>%
  ggplot(aes(y=Total_kBTUs, x=Date, fill= CUSTOMERCLASS))+
  geom_bar(stat='identity')
```

```{r}
pge_both %>% 
  filter(CUSTOMERCLASS %in% c('Gas- Commercial','Elec- Commercial','Gas- Residential','Elec- Residential'), YEAR %in% c(2019,2020,2021)) %>%
  ggplot(aes(y=cumsum(Total_kBTUs), x=Date, fill= CUSTOMERCLASS))+
  geom_bar(stat='identity')


```
